Multimodal spatial availability: a singly-constrained measure of accessibility considering multiple modes

### 2024-01-25 TRB-Poster.Rmd
Same changes to TRB poster

---
title: "Multimodal spatial availability: a singly-constrained measure of accessibility considering multiple modes"
poster_height: "121cm"
poster_width: "242cm"
font_family: 'Rasa'
primary_colour: "#7A003C"
secondary_colour: "#d3d3d3"
accent_colour: "#FDBF57"
titlebox_bordercol: "#5E6A71"
titlebox_borderwidth: "0.5cm"
author:
  - name: Anastasia Soukhov
    affil: 1
    email: soukhoa@mcmaster.ca
    twitter: soukhov_a
    main: true
  - name: Javier Tarriño-Ortiz
    affil: 2
    email: javier.tarrino.ortiz@upm.es
  - name: Julio A. Soria-Lara
    affil: 2
    email: julio.soria-lara@upm.es
  - name: Antonio Páez
    email: paezha@mcmaster.ca
    affil: 1
affiliation:
  - num: 1
    address: Department of Earth, Environment and Society, McMaster University, Hamilton, Canada
  - num: 2
    address: Centro de Investigación del Transporte (TRANSyT), Universidad Politécnica de Madrid, Madrid, Spain
column_numbers: 4
logoright_name: Presentation-images/acknowledgements-combo.png
logoleft_name: Presentation-images/soukhova-LinkTree.png
sectitle_textsize: "60pt"
sectitle_borderwidth: "1.5mm"
title_textsize: "110pt"
author_textsize: "80pt"
affiliation_textsize: "80pt"
body_textsize: "50px"
link-citations: true
bibliography: bibliography/bibliography.bib
output: 
  posterdown::posterdown_html:
    self_contained: true
knit: pagedown::chrome_print  #this knits the created html file into a .pdf. comment out if .pdf isn't needed
---
# Abstract
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r, eval=FALSE, include=FALSE}
# pagedown::chrome_print("TRB-Poster.Rmd", format = "pdf")
```
```{r load-packages, cache=FALSE, include=FALSE}
library(dplyr)
library(sf)
library(ggplot2)
library(tidyverse)
library(units)
library(patchwork)
library(ggspatial)
library(ggforce)
library(geomtextpath)
library(knitr)
library(gridExtra)
library(grid)
library(huxtable)

@@

Accessibility is a key concept in the analysis of land use and transportation systems. It is too coming of age from the perspective of planning [see _inter alia_, @handyAccessibility2020; @levinsonTransportAccessManual2020; @elGeneidyMaking2022]. Beginning with the work of Hansen [@hansenHowAccessibilityShapes1959], accessibility measures have been widely used to evaluate the efficiency of transportation systems when combined with the distribution of opportunities in space.

The most common form of accessibility measure is based on the gravity model, and recent research has focused on the way opportunities are counted (summed). In these calculations, the sums are not constrained, meaning the same opportunity can enter the sum for different origins. This unconstrained counting is a pitfall in conventional methods as it leads to values of accessibility that are difficult to interpret, especially when comparing results between multiple modes.

@@

We demonstrate spatial availability's use in the case of multiple modes or, more generally, heterogeneous population segments with distinct travel behaviours. We illustrate its features using a synthetic example, an empirical example of low emission zones in Madrid, Spain, and suggest future research directions in its use to evaluate policy interventions.

# Overview of multi-modal accessibility measures

@@

- So, one can calculate spatial availability per ...
  - **$i$**: ( $V_{i} = \sum_{m} V^m_{i}$ )
  - **mode**: ( $v^{m} = \sum_{m} \sum_{i} V^m_{i}$ )
  - **capita in $i$**: ( $v_{i} = \sum_{m} \frac{V_i^m}{P_{i^m}}$ )
  - Etc.

@@

The calculated Hansen (**S**), Shen-type (**a**) and Spatial Availability (**V**) assuming an impedance function of $f(c_{ij})^{\text{x}} = f(c_{ij})^{\text{z}} =exp(-0.1 \cdot c_{ij}^m)$ is shown in the table.

![](Presentation-images/toy-example.png){#id .class width=100%}
When competition is not considered, **S** values for the Suburban A and Urban B are the same. Does this equivalency make sense for the differently sized A and B? Further, if constraints are not incorporated (i.e., **S** and **a**), values are hard to interpret. The regional sums of **S** and **a** are meaningless.

@@

# Empirical example: quantifying multimodal access in Madrid

@@

![](Presentation-images/Fig2.png){#id .class width=100%}

![](Presentation-images/Fig5.png){#id .class width=95%}

@@

![](Presentation-images/Fig6.png){#id .class width=100%}

@@

![](Presentation-images/Fig7.png){#id .class width=95%}

>>>>

---
title: "Multimodal spatial availability: a singly-constrained measure of accessibility considering multiple modes"
poster_height: "36in"
poster_width: "54in"
font_family: 'Rasa'
primary_colour: "#7A003C"
secondary_colour: "#d3d3d3"
accent_colour: "#FDBF57"
titlebox_bordercol: "#5E6A71"
titlebox_borderwidth: "0.2cm"
author:
  - name: Anastasia Soukhov
    affil: 1
    email: soukhoa@mcmaster.ca
    twitter: soukhov_a
    main: true
  - name: Javier Tarriño-Ortiz
    affil: 2
    email: javier.tarrino.ortiz@upm.es
  - name: Julio A. Soria-Lara
    affil: 2
    email: julio.soria-lara@upm.es
  - name: Antonio Páez
    email: paezha@mcmaster.ca
    affil: 1
affiliation:
  - num: 1
    address: Department of Earth, Environment and Society, McMaster University, Hamilton, Canada
  - num: 2
    address: Centro de Investigación del Transporte (TRANSyT), Universidad Politécnica de Madrid, Madrid, Spain
column_numbers: 4
logoright_name: Presentation-images/acknowledgements-combo.png
logoleft_name: Presentation-images/soukhova-LinkTree.png
sectitle_textsize: "40pt"
sectitle_borderwidth: "0.4mm"
title_textsize: "80pt"
author_textsize: "50pt"
affiliation_textsize: "40pt"
body_textsize: "35px"
link-citations: true
bibliography: bibliography/bibliography.bib
output: 
  posterdown::posterdown_html:
    self_contained: true
knit: pagedown::chrome_print  #this knits the created html file into a .pdf. comment out if .pdf isn't needed
---
# Overview
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r, eval=FALSE, include=FALSE}
# pagedown::chrome_print("TRB-Poster.Rmd", format = "pdf")
```
```{r load-packages, cache=FALSE, include=FALSE}
library(dplyr)
library(sf)
library(ggplot2)
library(tidyverse)
library(units)
library(patchwork)
library(ggspatial)
library(ggforce)
library(geomtextpath)
library(knitr)
library(gridExtra)
library(grid)
library(huxtable)

@@

Accessibility is a key concept in the analysis of land use and transportation systems. It is too coming of age from the perspective of planning [see _inter alia_, @handyAccessibility2020; @elGeneidyMaking2022]. Beginning with the work of Hansen [@hansenHowAccessibilityShapes1959], accessibility measures have been widely used to evaluate the efficiency of transportation systems when combined with the distribution of opportunities in space.

The most common form of accessibility measure is based on the gravity model, and recent research has focused on the way opportunities are counted (summed). Notably, the sums are not constrained meaning the same opportunity can enter the sum for different origins. This unconstrained counting results in accessibility values that are difficult to interpret, especially when comparing results between multiple modes.

@@

We demonstrate spatial availability's use in the case of multiple modes (i.e., heterogeneous population segments with distinct travel behaviours). We illustrate its features using a synthetic example, an empirical example of low emission zones in Madrid, Spain, and suggest future research directions in its use to evaluate policy interventions.

# General form of multimodal accessibility measures

@@

- So, one can calculate spatial availability per ...
  - **$i$**: ( $V_{i} = \sum_{m} V^m_{i}$ )
  - **mode**: ( $v^{m} = \sum_{m} \sum_{i} V^m_{i}$ )
  - **capita in $i$**: ( $v_{i} = \sum_{m} \frac{V_i^m}{P_{i^m}}$ )

@@

![_Figure 1: Calculated Hansen $S$, Shen-type $a$ and Spatial Availability $V$ assuming an impedance function of $f(c_{ij})^{\text{x}} = f(c_{ij})^{\text{z}} =exp(-0.1 \cdot c_{ij}^m)$ for the multimodal synthetic population._](Presentation-images/toy-example.png){#id .class width=100%}

When competition is not considered, **S** values for the Suburban A and Urban B are the same. Does this equivalency make sense for the differently sized A and B? Further, if constraints are not incorporated (i.e., **S** and **a**), values are hard to interpret. The regional sums of **S** and **a** are meaningless.

@@

# Empirical example: multimodal access in Madrid

@@

![_Figure 2: Spatial distribution of jobs taken by people living and working in Madrid as reported in the 2018 travel survey (left) and trip length distribution (right). Grey TAZs have no jobs. Ranges of values in the legend are quintiles. The TAZ shapefile is available from the Community of Madrid open data portal._](Presentation-images/Fig2.png){#id .class width=100%}

![_Figure 3: Spatial availability of jobs per origin and mode $V_i^m$._](Presentation-images/Fig5.png){#id .class width=98%}

@@

![_Figure 4: Proportion of population by mode and spatial availability of jobs by mode ($V_i^m$) aggregated for the city of Madrid (All), the area within the M-30 highway, and within the Centro._](Presentation-images/Fig6.png){#id .class width=100%}

@@

![_Figure 5: Distribution of spatially available jobs per capita by mode of transportatin ($v_i^m$)._](Presentation-images/Fig7.png){#id .class width=98%}
